---
output:
  md_document:
    variant: markdown_github
---

# Gesis

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/gesis)](http://cran.r-project.org/package=gesis)
[![](http://cranlogs.r-pkg.org/badges/grand-total/gesis)](http://cran.r-project.org/web/packages/gesis)

## Introduction

The [GESIS Data Catalogue](https://dbk.gesis.org/) offers a repository of 
approximately 5,000 datasets. Due to a lack of an API, however, accessing these
datasets in a programmatic and reproducible way is difficult. The `gesis`
package seeks to solve this issue through the use of
[Selenium](http://www.seleniumhq.org/) and the `RSelenium` package. 

To install the package from github:

```{r options}
knitr::opts_chunk$set(eval = FALSE)
```


```{r}
# install.packages("devtools")
devtools::install_github("expersso/gesis")
```

In essence, the `gesis` package allows the user to emulate a web browser
session, wherein he or she logs in to the GESIS website, browses to the data set
of interest, clicks to download that dataset, agrees to accept the terms of use,
and, ultimately, downloads the dataset. This whole process can be done through
`R` with these three steps:

1. Initiating a Selenium server (`setup_gesis`)
1. Logging in to GESIS (`gesis_login`)
1. Downloading a specified dataset (`download_dataset`)

The next section describes this workflow by working through a simple example. An
second example then shows how the package can be leveraged for more advanced
uses.

## A simple example

The first step to using the `gesis` package is always to set up a Selenium 
remote driver by running `setup_gesis`. This function takes care of all the 
preliminaries for using Selenium, including checking for the existence of a 
Selenium server, starting such a server, creating a remote driver, and opening a
web browser window. In addition, it specifies browser settings such that files 
can be downloaded without prompting the user. (NB: The `setup_gesis` function
currently only supports Firefox. See the help file for how to use other
browsers.)

```{r setup}
if(!dir.exists("downloads")) dir.create("downloads")
gesis_remDr <- setup_gesis(download_dir = "downloads")
```

An empty browser window should now pop up. Leave this window open; this is where
we will emulate a session to access the GESIS website.

Next we go to the GESIS main page and log in by providing our user name and 
password. (To avoid having to provide the user name and password in plain text 
in a script, the default behavior is to fetch these as options using 
`getOption("gesis_user")` and `getOption("gesis_pass")`. You can thus specify 
these in your `.Rprofile` by `option("gesis_user" = "myusername", "gesis_pass" =
"mypassword")`.)

```{r}
login_gesis(gesis_remDr, user = "myusername", pass = "mypassword")
```

Switching to the browser window opened earlier, we should now see that we are
logged in. Now all we have to do is figure out the unique identifier for the
data set we are interested in. This is called a "DOI" and can be found on every
data set's description page.

```{r}
download_dataset(gesis_remDr, doi = 5928, filetype = "dta", purpose = 1)
```

The above function will:

1. direct the browser to go to the description page for the dataset with DOI 5928,
1. click to download the `.dta` (Stata) version of this dataset,
1. move to a separate download window that pops up,
1. specify that the dataset is to be used for scientific purposes,
1. accept the terms of use,
1. click "download",
1. accept the download without prompting the user,
1. close the pop-up window
1. move focus back to the first window, which is now ready for another download.

Finally, we can now check that the downloaded file is in the folder we
specified, and then close the browser window and the Selenium server.

```{r}
dir("downloads")
gesis_remDr$Close()
gesis_remDr$closeServer()
```

To simplify further analysis, the package also provides a convenience function 
for browsing the codebook of a specified dataset. This function does not require
an active Selenium session, but *does* require that the `xml2` package be
installed:

```{r}
browse_codebook(doi = 5928)
```

## A more realistic example

The workflow described above is clearly more laborious than just downloading
data sets by hand if you are only downloading a handful of data sets. However,
many opinion surveys take the form of repeated cross-sections, meaning that each
time a survey is conducted it is distributed as a separate file. If one is
interested in analyzing these surveys over time, one therefore needs to download
a separate data set for each point in time.

An example of such a repeated cross-section is a study called "Atlantic Trends",
for which there are annual surveys between 2002 and 2013. We can easily scrape
the DOI for these data sets.

```{r, eval = TRUE}
library(xml2)

# Browsing the gesis website, we find the url for the main page for these studies
url <- "https://dbk.gesis.org/dbksearch/GDesc2.asp?no=0074&ll=10&db=d&notabs=1"

page <- read_html(url)
doi_links <- xml_find_all(page, "//a[contains(text(), 'ZA')]")
doi <- substr(xml_text(doi_links), 3, 7)
str(doi)
```

Using the `gesis` package just like before, we can now batch download all these
surveys:

```{r}
# Setup preliminaries
if(!dir.exists("downloads")) dir.create("downloads")
gesis_remDr <- setup_gesis(download_dir = "downloads")

# Log in 
login_gesis(gesis_remDr, user = "myusername", pass = "mypassword")

# Loop over DOIs to download
lapply(doi, download_dataset, remDr = gesis_remDr)
```

Disclaimer: the `gesis` package is neither affiliated with, nor endorsed by, the
Leibniz Institute for the Social Sciences. I have been unable to find any 
indication that programmatic access to the website is disallowed under its terms
of use (indeed, its 
[guidelines](https://dbk.gesis.org/dbksearch/guidelines.asp) appear to 
encourage it). That said, I would discourage users from using the `gesis`
package to put undue pressure on their servers by initiating unnecessary (or
unnecessarily large) batch downloads.