---
title: "Programmatic access to the GESIS Data Catalogue (DBK)"
author: "Eric Persson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Programmatic access to the GESIS Data Catalogue (DBK)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The [GESIS Data Catalogue](https://dbk.gesis.org/) offers a repository of 
approximately 5,000 datasets. 

```{r load}
library(gesis)
```

We start by listing all available groups of studies:

```{r groups}
groups <- get_study_groups()
head(groups, 10)
```

We see that the Eurobarometer has study group number 0008. Let's looks at all
available Eurobarometer waves:

```{r eurobar_waves}
eurobars <- get_datasets("0008")
head(eurobars)
```

We would now like to download the first three studies. We first need to log in
to the Gesis website and then pass the DOIs (unique data set identifiers) to 
`download_dataset`:

```{r login, eval=FALSE}
# username and password stored as environment 
# variables "GESIS_USER" and "GESIS_PASS"
gesis_session <- login()
```

```{r create_dir}
if(!dir.exists("downloads")) dir.create("downloads")
```

```{r, eval=FALSE}
download_dataset(s = gesis_session, doi = eurobars$doi[1:3], 
                 path = "downloads", filetype = ".dta")
```

```{r list_files}
files <- list.files("downloads", full.names = TRUE)
```

We can also download the codebooks for the same studies:

```{r codebooks}
download_codebook(eurobars$doi[1:3], path = "downloads")
```

Using the `haven` package we can now read the data sets:

```{r read_data, eval=FALSE}
# library(haven)
df <- read_dta(files[1])
dim(df)
```

```{r remove_downloads, echo=FALSE}
unlink("downloads", recursive = TRUE)
```

Disclaimer: the `gesis` package is neither affiliated with, nor endorsed by, the
Leibniz Institute for the Social Sciences. I have been unable to find any 
indication that programmatic access to the website is disallowed under its terms
of use (indeed, its 
[guidelines](https://dbk.gesis.org/dbksearch/guidelines.asp) appear to 
encourage it). That said, I would discourage users from using the `gesis`
package to put undue pressure on their servers by initiating unnecessary (or
unnecessarily large) batch downloads.