---
title: "Programmatic access to the GESIS Data Catalogue (DBK)"
author: "Bill Clay"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Programmatic access to the GESIS Data Catalogue (DBK)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
The [GESIS Data Catalogue](https://dbk.gesis.org/) offers a repository of 
approximately 5,000 datasets. Due to a lack of an API, however, accessing these
datasets in a programmatic and reproducible way is difficult. The `gesis`
package seeks to solve this issue through the use of
[Selenium](http://www.seleniumhq.org/) and the `RSelenium` package. 

In essence, the `gesis` package allows the user to emulate a web browser
session, wherein he or she logs in to the GESIS website, browses to the dataset
of interest, clicks to download that dataset, agrees to accept the terms of use,
and, ultimately, downloads the dataset. This whole process can be done through
`R` with these three steps:

1. Initiating a Selenium server (`setup_gesis`)
1. Logging in to GESIS (`gesis_login`)
1. Downloading a specified dataset (`download_dataset`)

The next section describes this workflow by working through a simple example. An
extended example then shows how the package can be leveraged for more advanced
uses.

## A simple example

The first step to using the `gesis` package is always to set up a Selenium 
remote driver by running `setup_gesis`. This function takes care of all the 
preliminaries for using Selenium, including checking for the existence of a 
Selenium server, starting such a server, creating a remote driver, and opening a
web browser window. In addition, it specifies browser settings such that files
can be downloaded without prompting the user.

```{r setup, eval=FALSE}
if(!dir.exists("downloads")) dir.create("downloads")
gesis_remDr <- setup_gesis(download_dir = "downloads")
```

An empty browser window should now pop up. Leave this window open; this is where we will emulate a session to access the GESIS website.

Next we go to the GESIS main page and log in by providing our user name and
password. (To avoid having to provide the user name and password in plain text
in a script, the default behavior is to fetch these as options using
`getOption("gesis_user")` and `getOption("gesis_pass")`. You can thus specify
these in your `.Rprofile` by `option("gesis_user" = "myusername")` ...).

```{r, eval=FALSE}
login_gesis(gesis_remDr, user = "myusername", pass = "mypassword")
```

Switching to the browser window opened earlier, we should now see that we are logged in. Now all we have to do is figure out the unique identifier for the dataset we are interested in. This is called a "DOI" and can be found on every dataset's description page. 

```{r, eval=FALSE}
download_dataset(gesis_remDr, doi = "5928", filetype = "dta")
```

The above function will:

1. direct the browser to go to the description page for the dataset with DOI 5928,
1. click to download the `.dta` (Stata) version of this dataset,
1. move to a separate download window that pops up,
1. specify that the dataset is to be used for scientific purposes,
1. accept the terms of use,
1. click "download",
1. accept the download without prompting the user,
1. close the pop-up window
1. move focus back to the first window, which is now ready for another download.

Finally, we can now check that the downloaded file is in the folder we specified, and
then close the browser window and the Selenium server.

```{r, eval=FALSE}
dir("downloads")
gesis_remDr$Close()
gesis_remDr$closeServer()
```

To simplify further analysis, the package also provides a convenience function
for browsing the codebook of a specified dataset. This function does not require
an active Selenium session, but *does* require that packages `xml2` and `rvest`
be installed:

```{r, eval=FALSE}
browse_codebook(doi = "5928")
```

## An extended example