---
title: "gesis: Programmatic access to the GESIS Data Catalogue (DBK)"
author: "Eric Persson and Frederick Solt"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{icpsrdata: Reproducible Retrieval of Pew Research Center Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The [GESIS Data Catalogue](https://dbk.gesis.org/) offers a repository of 
approximately 5,000 datasets. Due to a lack of an API, however, accessing these
datasets in a programmatic and reproducible way is difficult. The `gesis`
package seeks to solve this issue through the use of
[Selenium](http://www.seleniumhq.org/) and the `RSelenium` package.  


## Setup

`gesis` depends on the Firefox browser; if it is not installed, [get it here](https://www.mozilla.org/firefox).

An optional, but recommended, setup step is to add the information GESIS requires to your [.Rprofile](http://www.statmethods.net/interface/customizing.html) as in the following example:

```{r eval = FALSE}
options("gesis_user" = "juanita-herrera@uppermidwest.edu",
        "gesis_pass" = "password123!")
```

The `gesis_download` function will then access the information it needs to pass on to GESIS by default.  This means that researchers will not have to expose their info in their R scripts and that others reproducing their results later will be able to execute those R scripts without modification.  (They will, however, need to enter their own information into their own .Rprofiles, a detail that should be noted in the reproducibility materials to avoid confusion.)


## Use

The `gesis_download` function 
1. opens a Firefox browser pointed to the GESIS Data Catalogue,
1. enters the required login information,
1. navigates to the GESIS page of specified dataset,
1. downloads the dataset's codebook, if one is available,
1. initiates a download of the dataset in the specified format,
1. indicates a specified purpose (by default, "scientific purposes") and agrees to the terms and conditions of the download,
1. downloads the dataset's files to a eponymous subdirectory of a specified directory (or, by default, a `gesis_data` directory in the current working directory).

Datasets are specified using the `doi` argument.  GESIS uses a DOI, that is, a unique number, to identify each of its datasets; this number is consistently listed alongside the dataset's name.  For the [2013 Transatlantic Trends Survey](https://dbk.gesis.org/dbksearch/SDesc2.asp?ll=10&notabs=1&af=&nf=&search=&search2=&db=E&no=5646), for example, the file id is 5646.

To reproducibly download this dataset, then:

```{r eval=FALSE}
gesis_download(doi = 5646)
```


Many opinion surveys, however, take the form of repeated cross-sections, meaning that each time a survey is conducted it is distributed as a separate file. If one is interested in analyzing these surveys over time, one therefore needs to download a separate data set for each point in time. Transatlantic Trends is one such repeated cross-sectional study: there are annual surveys between 2002 and 2013. We can easily scrape the DOI for these data sets.

```{r eval=FALSE}
library(xml2)

# Browsing the gesis website, we find the url for the main page for these studies
tt_url <- "https://dbk.gesis.org/dbksearch/GDesc2.asp?no=0074&ll=10&db=d&notabs=1"

tt_page <- read_html(url)
tt_doi_links <- xml_find_all(page, "//a[contains(text(), 'ZA')]")
tt_doi <- substr(xml_text(tt_doi_links), 3, 7)
```

All of these datasets may be downloaded in a single command by passing this vector of DOI to `doi`:

```{r eval=FALSE}
gesis_download(doi = tt_doi)
```

Each of these data sets and their respective codebooks are then found in subdirectories of the `gesis_data` directory within the current working directory; note that this default location may be changed using the `download_dir` argument. They are then ready to be loaded into R using, e.g., `rio::import`.

Disclaimer: the `gesis` package is neither affiliated with, nor endorsed by, the Leibniz Institute for the Social Sciences. I have been unable to find any indication that programmatic access to the website is disallowed under its terms of use (indeed, its 
[guideslines](https://dbk.gesis.org/dbksearch/guidelines.asp) appear to encourage it). That said, I would discourage users from using the `gesis` package to put undue pressure on their servers by initiating unnecessary (or unnecessarily large) batch downloads.
